---
title: "506_final_project_EDA"
output: word_document
date: "2022-12-02"
---

```{r}
library(fpp2)
library(readr)
library(forecast)
library(zoo)
library(AICcmodavg)
```
```{r}
car_sales <- read_csv("C:\\Users\\maha.jayapal\\Downloads\\monthly-car-sales.csv",show_col_types = FALSE)
#change the name of column
colnames(car_sales) <- c("Month", "Car_Sales")
View(car_sales)
```

```{r}
car_sales.ts <- ts(car_sales$Car_Sales, start=c(1960,1), end=c(1968,12), frequency=12)
```
```{r}
autoplot(car_sales.ts)
```
Random Walk:
```{r}
arima_1 <- arima(car_sales.ts, order=c(1,0,0))
arima_1
```
```{r}
acf(diff(car_sales.ts, lag=1), main="ACF plot for lag 1 series")
```
The AR 1 Coefficient is 0.7320 is closer to 1, but the autocorrelations of the
differenced series is not closer to zero, with all bars not within the
threshold this infers that it is not random.

```{r}
#ACF plot
acf(car_sales.ts,lag.max = 36)
```
There is strong positive autocorrelation at lags 12, 24, and 36 in monthly data wich will reflect an annual seasonality. It means values during a given month each year are positively correlated. 
```{r}
pacf(car_sales.ts, lag.max = 36)
```
The PACF shows that there is a seasonal lag at 12.

```{r}
car_sales.ts
```




